---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
keep_md: true
---

This is a R Markdown document.

#### Load all the library first
```{r, echo = TRUE, results='hide'}
library(dplyr, warn.conflicts = FALSE)
library(lubridate, warn.conflicts = FALSE)
library(ggplot2, warn.conflicts = FALSE)
```
## Loading and preprocessing the data

1. Load the data 
2. Process/transform the data (if necessary) into a format suitable for your analysis

```{r step1, echo = TRUE}
activity <- read.csv("./activity.csv", header = TRUE, sep = ",")
activity <- tbl_df(activity)
activity$date <- as.Date(as.character(activity$date), "%Y-%m-%d")
```

## What is mean total number of steps taken per day?

```{r step2, echo = TRUE}
dates <- as.factor(activity$date)
startDate <- as.Date("2012-10-01")
total <- activity %>%
    group_by(date) %>%
    summarize(t = sum(steps, na.rm = TRUE)) %>%
    mutate(days = as.numeric(date - startDate) + 1)
totalCount <- vector()
for (i in 1:nrow(total)) {
    totalCount <- c(totalCount, rep(total[[i, "days"]], total[[i, "t"]]))
}
```
Make a histogram of the total number of steps taken each day
```{r, echo = TRUE}
hist(totalCount, breaks = nrow(total), xlab = "Day", main = "Everyday Total Steps")
```

Calculate and report the mean and median of the total number of steps taken per day
```{r, echo = TRUE}
total1 <- activity %>%
    subset(!is.na(steps)) %>%
    group_by(date) %>%
    summarize(mean = mean(steps), median = median(steps))
head(total1)
```

## What is the average daily activity pattern?

Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r step3, echo = TRUE}
activity1 <- activity %>%
    mutate(datetime = `minute<-`(date, interval)) %>%
    filter(!is.na(steps))
with(activity1, plot(datetime, steps, type = "l", 
                     xlab = "Datetime", ylab = "Steps",
                     main = "Every Five Minutes Steps"))
abline(v = activity1[[max(activity1$steps, na.rm = TRUE), 'datetime']], 
       col = "red", lwd = 4)
```

## Imputing missing values

Calculate and report the total number of missing values in the dataset
```{r step4, echo = TRUE}
summary(is.na(activity$steps))
```

Filled the missing data
```{r, echo = TRUE}
dayMeans <- activity %>%
    group_by(date) %>%
    summarize(m = mean(steps, na.rm = TRUE))
dayMeans[is.nan(dayMeans$m), 'm'] <- 0
activityMissing <- activity
activityMissing$steps <- apply(as.matrix(activityMissing), 1, function(x) {
    if (is.na(x[1])) {
        dayMeans[dayMeans$date == x[2], 'm'][[1, 'm']]
    } else {
        as.numeric(x[1])
    }
})
```

Make a histogram of the total number of steps taken each day 
```{r, echo = TRUE}
total <- activityMissing %>%
    group_by(date) %>%
    summarize(t = sum(steps, na.rm = TRUE)) %>%
    mutate(days = as.numeric(date - startDate) + 1)
totalCount <- vector()
for (i in 1:nrow(total)) {
    totalCount <- c(totalCount, rep(total[[i, "days"]], total[[i, "t"]]))
}
hist(totalCount, breaks = nrow(total), xlab = "Day", main = "Filled Missing Data, Everyday Total Steps")
```

Calculate and report the mean and median total number of steps taken per day
```{r, echo = TRUE}
total1 <- activityMissing %>%
    group_by(date) %>%
    summarize(mean = mean(steps), median = median(steps))
head(total1)
```

## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day
```{r step5, echo = TRUE}
allWeekdays <- c(2:6)
activityMissing$w <- factor(wday(activityMissing$date) %in% allWeekdays,
                            level = c(FALSE, TRUE), labels = c("weekend", "weekday"))
activityMissing <- activityMissing %>%
    group_by(w) %>%
    mutate(x = 1:n())
```

Make a panel plot containing a time series plot
```{r, echo = TRUE}
ggplot(activityMissing, aes(x, steps)) +
    geom_line(color = "blue") +
    facet_wrap(facets = vars(w), nrow = 2) +
    xlab("Interval") +
    ylab("Steps") +
    theme(strip.background = element_rect(fill = "orange"), 
          panel.spacing = unit(0, "lines"))

```
