---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
keep_md: true
---

This is a R Markdown document.

#### Load all the library first
```{r, echo = TRUE, results='hide'}
library(dplyr, warn.conflicts = FALSE)
library(lubridate, warn.conflicts = FALSE)
library(ggplot2, warn.conflicts = FALSE)
```
## Loading and preprocessing the data

1. Load the data 
2. Process/transform the data (if necessary) into a format suitable for your analysis

```{r step1, echo = TRUE}
activity <- read.csv("./activity.csv", header = TRUE, sep = ",")
activity <- tbl_df(activity)
activity$date <- as.Date(as.character(activity$date), "%Y-%m-%d")
```

## What is mean total number of steps taken per day?

```{r step2, echo = TRUE}
total <- tapply(activity$steps, activity$date, sum, na.rm = TRUE)
```
Make a histogram of the total number of steps taken each day
```{r, echo = TRUE}
hist(total, breaks = 30, xlab = "Steps", main = "Histogram of Everyday Total Steps")
```

Calculate and report the mean and median of the total number of steps taken per day
```{r, echo = TRUE}
summary(total)
```

## What is the average daily activity pattern?

Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r step3, echo = TRUE}
activityAvg <- activity %>% 
  group_by(interval) %>%
  summarize(steps = mean(steps, na.rm = TRUE))
with(activityAvg, plot(interval, steps, type = "l", 
                     xlab = "Interval", ylab = "Steps",
                     main = "Interval Steps"))
abline(v = activityAvg[[which.max(activityAvg$steps), 'interval']], 
       col = "red", lwd = 4)

activityAvg[[which.max(activityAvg$steps), 'interval']]
```

## Imputing missing values

Calculate and report the total number of missing values in the dataset
```{r step4, echo = TRUE}
summary(is.na(activity$steps))
```

Filled the missing data
```{r, echo = TRUE}
dayMeans <- activity %>%
    group_by(date) %>%
    summarize(m = mean(steps, na.rm = TRUE))
dayMeans[is.nan(dayMeans$m), 'm'] <- 0
activityMissing <- activity
activityMissing$steps <- apply(as.matrix(activity), 1, function(x) {
    if (is.na(x[1])) {
        activityAvg[activityAvg$interval == as.integer(x[3]), 'steps'][[1, 'steps']]
    } else {
        as.integer(x[1])
    }
})
```

Make a histogram of the total number of steps taken each day 
```{r, echo = TRUE}
total <- tapply(activityMissing$steps, activityMissing$date, sum)
```
Make a histogram of the total number of steps taken each day
```{r, echo = TRUE}
hist(total, breaks = 30, xlab = "Steps", main = "Filled Missing Data, Histogram of Everyday Total Steps")
```

Calculate and report the mean and median of the total number of steps taken per day after filling missing data
```{r, echo = TRUE}
summary(total)
```

## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day
```{r step5, echo = TRUE}
activityMissing$w <- factor(wday(activityMissing$date) %in% allWeekdays,
                            level = c(FALSE, TRUE), labels = c("weekend", "weekday"))
activityMissingWeek <- activityMissing %>%
    group_by(w, interval) %>%
    summarize(steps = mean(steps))
```

Make a panel plot containing a time series plot
```{r, echo = TRUE}
ggplot(activityMissingWeek, aes(interval, steps)) +
    geom_line(color = "blue") +
    facet_wrap(facets = vars(w), nrow = 2) +
    xlab("Interval") +
    ylab("Steps") +
    theme(strip.background = element_rect(fill = "orange"), 
          panel.spacing = unit(0, "lines"))

```
